---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cbbdata

<!-- badges: start -->
<!-- badges: end -->

Welcome to `cbbdata`, an R package designed to grant swift and effortless access to the `cbbdata` API, offering an abundance of college basketball statistics and data! By harnessing the power of back-end SQL queries and direct file transfers, this package equips you with the essential information required to dissect and comprehend the intricacies of college basketball.

Getting started with cbbdata is a breeze – simply sign up for a free API key and delve into the rich world of college basketball data. With its intuitive interface and robust back-end capabilities, this package stands as an indispensable asset for anyone eager to explore the depths of college basketball analysis.

## Installation

You can install the development version of cbbdata from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("andreweatherman/cbbdata")
```

## Registering for an API key

An API key is free and easy to obtain by using the official `cbbdata` R package. To register, simply pass a username and password to the `cbd_create_account` endpoint. As of April 24, 2023, there is no way to retrieve a forgotten password.

```{r register, eval = FALSE}
# to register
cbbdata::cbd_create_account(username = 'xxx', password = 'xxx', confirm_password = 'xxx')
```

To obtain your key, pass the same credentials to the `cbd_login` endpoint. Your API key will then be stored as an environmental variable in your R session.

```{r login, eval = FALSE}
# to login
cbbdata::cbd_login(username = 'xxx', password = 'xxx')
```

## Example

...

## Elevated/Premium Access

Public access to the `cbbdata` API is completely free but hosting it isn't. If you want to support the project beyond sharing it, please consider donating below.


<Insert link here>
