---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%"
)
library(tidyverse)
```

# CBBData, Your Gateway to In-Depth College Basketball Insights

<!-- badges: start -->
<!-- badges: end -->

Getting reliable, up-to-date college basketball data shouldn't be a hassle. With `cbbdata`, it isn't. This package makes it straightforward to access a wide range of stats with minimal effort.

**Easy Access:** `cbbdata` simplifies the process of fetching college basketball statistics. No more scraping or searching through multiple sources: just clean, easy access to the data you need.

**Fast and Accurate:** Built with Flask and Python, `cbbdata` uses SQL queries and direct file transfers to ensure you get accurate and current data quickly. Our database is updated every 15 minutes during the season.

**Comprehensive Data:** Whether you need player stats, team analytics, game results, or advanced metrics, `cbbdata` has you covered with nearly 30 endpoints.

## Installation

You can install the development version of `cbbdata` from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("andreweatherman/cbbdata")
```

## Registering for an API key

An API key is free and easy to obtain by using the official `cbbdata` R package. To register, simply pass a username and password to the `cbd_create_account` endpoint. Your API key will be emailed to you – make sure to check your spam folder -- but you need not manually store your API key anywhere. The preferred way to access your account is with your username and password.

*Note: If you want to access KenPom data, your CBBData account email must match your KenPom account email*

```{r register, eval = FALSE}
# to register
cbbdata::cbd_create_account(username = 'xxx', email = 'xxx', password = 'xxx', confirm_password = 'xxx')
```

## Obtaining Your Key

After registering, there are two ways to obtain your key using the `cbbdata` R package.

### By-Session Log-In:

To obtain your key for use in your current R session, you can pass your username and password to the `cbd_login` function. This will retrieve your key and store it as a session variable. If you restart or leave your R session, you will need to log-in again.

```{r one_off_login, eval = FALSE}
# per-session log-in
cbbdata::cbd_login(username = 'xxx', password = 'xxx')
```

### Persistent Log-In (Preferred):

The recommended way to interact with the `CBBData` API is to store your username (`CBD_USER`) and password (`CBD_PW`) inside the .Renviron file. If you are unsure on how to do this, the `cbd_login` function will walk you through the process. Please note that this *will* require restarting your R session.

```{r persistent_login, eval = FALSE}
# persistent log-in
cbbdata::cbd_login()
```

If your API key is not stored as a session variable, `cbbdata` functions will grab your credentials from the .Renviron file and automatically log you in. With this method, you will not have to log-in again.

## Available Data

`cbbdata` provides broad access to leading college basketball resources. `cbbdata` is continuously growing and the available data includes:

### Barttorvik:

Developer [Andrew Weatherman](https://twitter.com/andreweatherman) wrote the popular `toRvik` R package. `cbbdata` replaces `toRvik` and brings with it a rich collection of Barttorvik data.

#### Metric Ratings:
-   Year-end ratings (`cbd_torvik_ratings`)
-   Day-by-day ratings (`cbd_torvik_ratings_archive`)
-   Team four factor splits (`cbd_torvik_team_factors`)

E.g., if you want to see what the no-bias T-Rank top 10 looks like:
```{r, message=FALSE, warning=FALSE}
cbbdata::cbd_torvik_team_factors(year = 2024, no_bias = TRUE) %>%
  dplyr::slice(1:10) %>%
  dplyr::select(team, barthag, adj_o, adj_d)
```

#### Player Data:
-   Individual game logs (`cbd_torvik_player_game`)
-   Season averages (`cbd_torvik_player_season`)
-   Season splits (`cbd_torvik_player_split`)

E.g., if you want to see which ACC player averages the most points at home:
```{r, message=FALSE, warning=FALSE}
cbbdata::cbd_torvik_player_split(year = 2024, conf = 'ACC', split = 'location') %>%
  dplyr::filter(games >= 10) %>%
  dplyr::slice_max(pts, n = 1) %>%
  dplyr::select(team, player, pts, games)
```

#### Team + Conference Data:
-   Team stats splits (`cbd_torvik_team_split`)
-   Team histories (`cbd_torvik_team_history`)
-   Conference four factor splits (`cbd_torvik_conf_factors`)

E.g., if you want to see which conferences shoot the best at home against top 100 teams:
```{r, message=FALSE, warning=FALSE}
cbbdata::cbd_torvik_conf_factors(2024, venue = 'home', top = 100) %>%
  dplyr::filter(games >= 5) %>%
  dplyr::slice_max(efg, n = 5) %>%
  dplyr::select(conf, games, efg)
```

#### Game Data:
-   Individual game box (`cbd_torvik_game_box`)
-   Individual game four factors (`cbd_torvik_game_factors`)
-   Individual game stats (box + factors) (`cbd_torvik_game_stats`)
-   Season schedule (`cbd_torvik_season_schedule`)

E.g., if you want to track how Duke's offense has performed this season:
```{r, message=FALSE, warning=FALSE}
cbbdata::cbd_torvik_game_factors(year = 2024, team = 'Duke') %>%
  ggplot2::ggplot(aes(date, adj_o)) +
  ggplot2::geom_line() +
  ggplot2::geom_point(aes(color = result), size = 3) +
  ggplot2::scale_color_manual(values = c('darkred', 'darkgreen'), guide = NULL) +
  ggplot2::theme_minimal()
```

#### Predictions
-   Individual game predictions (`cbd_torvik_game_prediction`)
-   Team season predictions (`cbd_torvik_season_prediction`)
-   Team season simulations (`cbd_torvik_season_simulation`)

E.g., if you want to run 10,000 simulations of Duke's season with their performance as of today:
```{r, message=FALSE, warning=FALSE}
# no date -> defaults to today
cbbdata::cbd_torvik_season_simulation('Duke', 2024) %>%
  ggplot2::ggplot(aes(wins, probability)) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::label_percent())
```

#### Tournament Results + Resumes
-   Daily NET rankings and quadrant records (`cbd_torvik_current_resume`)
-   Historical NET rankings and per-day metrics (`cbd_net_archive`)
-   Tournament performance (`cbd_torvik_ncaa_results`)
-   Tournament "committee sheets" (`cbd_torvik_ncaa_sheets`)
-   Resume database (`cbd_torvik_resume_database`)
-   Similar team tournament resumes (`cbd_torvik_similar_resumes`)

E.g., if you want to pull the five teams with the most Q1 NET wins:
```{r, message=FALSE, warning=FALSE}
cbbdata::cbd_torvik_current_resume() %>%
  dplyr::mutate(q1_wins = readr::parse_number(quad1)) %>%
  dplyr::slice_max(q1_wins, n = 5) %>%
  dplyr::select(team, conf, q1_wins, net)
```

### KenPom

To access KenPom data, you must have an active KenPom subscription and your CBBData account email must match your KenPom account email; this exists to curb account sharing. To activate your account, pass your **KenPom** account password through the `cbd_kenpom_authorization` function. The CBBData API will then confirm your account is active and log your expiration date. You will only be asked to re-authorize on that date.

```{r kenpom, eval = FALSE}
# persistent log-in
cbbdata::cbd_kenpom_authorization(password = 'xxx')
```

#### Metric Ratings: 
- Year-end ratings (`cbd_kenpom_ratings`)
- Day-by-day ratings (`cbd_kenpom_ratings_archive`)

